---
title: "Water Chemistry Analysis"
author: "Samuel Dumas"
date: "Winter Semester 2024"
output: pdf_document
---

```{r}
library(tidyverse)
library(dplyr)
library(lme4)
library(lmerTest)
library(ggpubr)
library(vegan)
library(gtools)
library(car)
```
# Global

```{r}
PONDS <- c("T106A1", "T106A2", "T106A3", "T106B1", "T106B2", "T106B3",
           "T108A1", "T108A2", "T108A3", "T108B1", "T108B2", "T108B3",
           "T109A1", "T109A2", "T109A3", "T109B1", "T109B2", "T109B3")
```


# Water Chemistry

```{r}
verticalData <- read_csv("Urban_Stream_2023_Site_Sample_Data_VP.csv")
```

```{r}
head(verticalData)
```

```{r}
pondIDs <- c("T106", "T108", "T109")
verticalSubData <- verticalData %>% 
  subset(Pond %in% pondIDs & Depth < 160)

numLabs <- c("July", "August")
names(numLabs) <- c("1", "2")
```


## GRAPHS

```{r Temperature}
verticalSubData %>% 
  ggplot(aes(x = Depth, y = Temperature, colour = Pond, shape = factor(Num))) +
  geom_point()  +
  geom_line(data = subset(verticalSubData, verticalSubData$Num == 1), aes(x = Depth, y = Temperature, group = Pond, colour = Pond)) +
  geom_line(data = subset(verticalSubData, verticalSubData$Num == 2), aes(x = Depth, y = Temperature, group = Pond, colour = Pond)) +
  coord_flip() + 
  scale_y_reverse() +
  scale_x_reverse() +
  scale_shape_manual("Month", values = c(1,2), labels = c("July", "August")) +
  labs(title = "Temperature versus Depth (July and August)", x = "Depth (cm)", y = "Temperature (\u00B0C)",
       colour = "Pond ID", shape = "Month")
ggsave("Plots/temp.png")
```

```{r}
verticalSubData %>% 
  ggplot(aes(x = Depth, y = DO, colour = Pond, shape = factor(Num))) +
  geom_point()  +
  geom_line(data = subset(verticalSubData, verticalSubData$Num == 1), aes(x = Depth, y = DO, group = Pond, colour = Pond)) +
  geom_line(data = subset(verticalSubData, verticalSubData$Num == 2), aes(x = Depth, y = DO, group = Pond, colour = Pond)) +
  coord_flip() + 
  scale_y_reverse() +
  scale_x_reverse() +
  scale_shape_manual("Month", values = c(1,2), labels = c("July", "August")) +
  labs(title = "Dissolved versus Depth (July and August)", x = "Depth (cm)", y = "Dissolved Oxygen (mg/L)",
       colour = "Pond ID")
ggsave("Plots/DO.png")
```

```{r}
verticalSubData %>% 
  ggplot(aes(x = Depth, y = Conductivity, colour = Pond, shape = factor(Num))) +
  geom_point()  +
  geom_line(data = subset(verticalSubData, verticalSubData$Num == 1), aes(x = Depth, y = Conductivity, group = Pond, colour = Pond)) +
  geom_line(data = subset(verticalSubData, verticalSubData$Num == 2), aes(x = Depth, y = Conductivity, group = Pond, colour = Pond)) +
  coord_flip() + 
  scale_x_reverse() +
  scale_shape_manual("Month", values = c(1,2), labels = c("July", "August")) +
  labs(title = "Conductivity versus Depth (July and August)", x = "Depth (cm)", y = "Conductivity (\u00b5s/cm)",
       colour = "Pond ID")
ggsave("Plots/cond.png")
```

```{r}
verticalSubData %>% 
  ggplot(aes(x = Depth, y = Turbidity, colour = Pond, shape = factor(Num))) +
  geom_point()  +
  geom_line(data = subset(verticalSubData, verticalSubData$Num == 1), aes(x = Depth, y = Turbidity, group = Pond, colour = Pond)) +
  geom_line(data = subset(verticalSubData, verticalSubData$Num == 2), aes(x = Depth, y = Turbidity, group = Pond, colour = Pond)) +
  coord_flip() + 
  scale_x_reverse() +
  scale_shape_manual("Month", values = c(1,2), labels = c("July", "August")) +
  labs(title = "Turbidity versus Depth (July and August)", x = "Depth (cm)", y = "Turbidity (NTU)",
       colour = "Pond ID")
ggsave("Plots/turb.png")
```

## LINEAR MODELS

```{r}
tempDepthModel1 <- lm(Temperature ~ Depth, verticalSubData)
summary(tempDepthModel1)

tempDepthModel <- function(pondID,
                           num){
  tempModel <- lm(Temperature ~ Depth, subset(verticalSubData, Pond == pondID & Num == num))
  return(summary(tempModel))
}

tempDepthModel('T106', 1)
tempDepthModel('T108', 1)
tempDepthModel('T109', 1)
tempDepthModel('T106', 2)
tempDepthModel('T108', 2)
tempDepthModel('T109', 2)
```

```{r}
doDepthModel1 <- lm(DO ~ Depth, verticalSubData)
summary(doDepthModel1)

doDepthModel <- function(pondID,
                           num){
  tempModel <- lm(DO ~ Depth, subset(verticalSubData, Pond == pondID & Num == num))
  return(summary(tempModel))
}

doDepthModel('T106', 1)
doDepthModel('T108', 1)
doDepthModel('T109', 1)
doDepthModel('T106', 2)
doDepthModel('T108', 2)
doDepthModel('T109', 2)
```

```{r}
turbDepthModel1 <- lm(Turbidity ~ Depth, verticalSubData)
summary(turbDepthModel1)

turbDepthModel <- function(pondID,
                           num){
  tempModel <- lm(Turbidity ~ Depth, subset(verticalSubData, Pond == pondID & Num == num))
  return(summary(tempModel))
}

turbDepthModel('T106', 1)
turbDepthModel('T108', 1)
turbDepthModel('T109', 1)
turbDepthModel('T106', 2)
turbDepthModel('T108', 2)
turbDepthModel('T109', 2)
```

# Invertebrates Data

```{r}
invertData <- read.csv("EEB397_InvertData.csv", row.names = 1)
```


```{r}
rank.abund <- function(x, num) {
  tmp <- x[num, -c(1,2)]
  tmp2 <- tmp[tmp > 0] / sum(tmp)
  tmp3 <- tmp2[order(-tmp2)]
cbind(rank = 1:length(tmp3), RA = tmp3) %>% 
  as.data.frame() %>% 
ggplot(aes(x = rank, y = RA)) +
  geom_col(colour ="blue", fill = "lightblue") +
  labs(y = "Relative Abundance", x = "Rank", title = paste("Pond", PONDS[num])) +
  ylim(0,1)
}
```

```{r}
A1 <- rank.abund(invertData, 1)
A2 <- rank.abund(invertData, 2)
A3 <- rank.abund(invertData, 3)
B1 <- rank.abund(invertData, 4)
B2 <- rank.abund(invertData, 5)
B3 <- rank.abund(invertData, 6)
A4 <- rank.abund(invertData, 7)
A5 <- rank.abund(invertData, 8)
A6 <- rank.abund(invertData, 9)
B4 <- rank.abund(invertData, 10)
B5 <- rank.abund(invertData, 11)
B6 <- rank.abund(invertData, 12)
A7 <- rank.abund(invertData, 13)
A8 <- rank.abund(invertData, 14)
A9 <- rank.abund(invertData, 15)
B7 <- rank.abund(invertData, 16)
B8 <- rank.abund(invertData, 17)
B9 <- rank.abund(invertData, 18)
```

```{r}
ggarrange(A1, A2, A3, ncol = 3) + 
  theme(plot.caption = element_text(hjust = 0)) +
  labs(caption = "Figure 2A. Rank-abundance distributions for the pond T106 site A.
")
ggsave("Plots/rankdistT106A.png")

ggarrange(A4, A5, A6, ncol = 3) + 
  theme(plot.caption = element_text(hjust = 0)) +
  labs(caption = "Figure 2B. Rank-abundance distributions for the pond T108 site A.
")
ggsave("Plots/rankdistT108A.png")

ggarrange(A7, A8, A9, ncol = 3) + 
  theme(plot.caption = element_text(hjust = 0)) +
  labs(caption = "Figure 2C. Rank-abundance distributions for the pond T109 site A.
")
ggsave("Plots/rankdistT109A.png")


ggarrange(B1, B2, B3, ncol = 3) + 
  theme(plot.caption = element_text(hjust = 0)) +
  labs(caption = "Figure 2D. Rank-abundance distributions for the pond T106 site B.
")
ggsave("Plots/rankdistT106B.png")

ggarrange(B4, B5, B6, ncol = 3) + 
  theme(plot.caption = element_text(hjust = 0)) +
  labs(caption = "Figure 2E. Rank-abundance distributions for the pond T108 site B.
")
ggsave("Plots/rankdistT108B.png")

ggarrange(B7, B8, B9, ncol = 3) + 
  theme(plot.caption = element_text(hjust = 0)) +
  labs(caption = "Figure 2F. Rank-abundance distributions for the pond T109 site B.
")
ggsave("Plots/rankdistT109B.png")
```


```{r}
invertData %>% 
  mutate(nSpecies=rowSums(.!=0) - 2) %>% 
  ggplot(aes(x = nSpecies)) +
  geom_bar(colour = "blue", fill = "lightblue") +
  theme(plot.caption = element_text(hjust = 0)) +
  labs(x = "Number of Species", y = "Frequency", title = "Frequency versus Number of Species",
       caption = "Figure 3A. Frequency distribution of number of species found in each sample (n = 18).")
ggsave("Plots/speciesdistribution.png")
```

```{r}
invertData %>% 
  mutate(nSpecies=rowSums(.!=0) - 2) %>% 
  ggplot(aes(x= nSpecies, fill = factor(depth), group = depth)) + 
    geom_bar(colour = "black", position = "stack") +
  theme(plot.caption = element_text(hjust = 0)) +
  labs(x = "Number of Species", y = "Frequency", 
       fill = "Depth (cm)",
       title = "Frequency versus Number of Species",
       caption = "Figure 3B. Frequency distribution of number of species found in each sample (n = 18), separated by depth.")
ggsave("Plots/speciesdistributionD.png")
```



```{r}
invertData %>% 
  pivot_longer(!c(pondID, depth), values_to = "total", names_to = "family") %>% 
  ggplot(aes(x=family, y = total, fill=family)) + 
         geom_bar(stat = "identity", position = "fill") + 
    theme(axis.text.x=element_blank(), 
          axis.ticks.x=element_blank(),
          axis.text.y=element_blank(), 
          axis.ticks.y=element_blank(),
          plot.caption = element_text(hjust = 0)) + 
  labs(x = "Present Taxonomic Families", y="", title = "Species Richness across Ponds.",
       fill = "Taxonomic Families",
       caption = "Figure 4A. Taxonomic families present in each sample.") + 
  guides(fill=guide_legend(ncol=2)) + 
         facet_wrap(vars(pondID), nrow = 6, ncol =3)
ggsave("Plots/familypresence.png")
```


```{r}
invertData %>% 
  pivot_longer(!c(pondID, depth), values_to = "total", names_to = "family") %>% 
  ggplot(aes(x=pondID, y = total, fill=family)) + 
         geom_bar(stat = "identity", position = "fill") + 
    theme(axis.text.x=element_text(angle = 45, hjust = 1),
          plot.caption = element_text(hjust = 0)) + 
  labs(x = "Present Taxonomic Families", y="Relative Abundance", title = "Species Richness across Ponds.",
       fill = "Taxonomic Families",
       caption = "Figure 4B. Relative abundance of taxonomic families by sample.") + 
  guides(fill=guide_legend(ncol=2))
ggsave("Plots/familyrelative.png")
```

```{r}
SR <- specnumber(invertData[,-c(1,2)])
simpInv <- diversity(invertData[,-c(1,2)], index = "invsimpson")
simpEven <- simpInv / SR
```


```{r}
brayCurtisInv <- 1 - vegdist(invertData[,-c(1,2)], method = "bray", diag = FALSE) %>% 
  as.matrix()
```


## Bray Curtis By Depth

```{r}
mod3index1 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(1, 18, 3)) %>% 
  as.data.frame()

mod3index2 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(2, 18, 3)) %>% 
  as.data.frame()

mod3index3 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(3, 18, 3)) %>% 
  as.data.frame()
```


```{r}
brayByDepth <- c()
for (i in 1:length(mod3index1$V1)){
  brayByDepth <- rbind(brayByDepth, 
                       c(50, brayCurtisInv[mod3index1$V1[i], mod3index1$V2[i]]))
}


temp2 <- 0 
for (i in 1:length(mod3index2$V1)){
  brayByDepth <- rbind(brayByDepth, 
                       c(100, brayCurtisInv[mod3index2$V1[i], mod3index2$V2[i]]))
}


temp3 <- 0 
for (i in 1:length(mod3index3$V1)){
  brayByDepth <- rbind(brayByDepth, 
                       c(150, brayCurtisInv[mod3index3$V1[i], mod3index3$V2[i]]))
}

brayByDepth <- as.data.frame(brayByDepth)
colnames(brayByDepth) <- c("depth", "bcSim")
```


```{r}
brayByDepth %>% 
  ggplot(aes(x = depth, y = bcSim, group = depth)) +
  geom_boxplot() + 
  labs(x = "Depth (cm)", y = "Similarity (Bray-Curtis)", title = "Bray-Curtis Similarity versus Depth")
```


## Bray-Curtis By Pond


```{r}
ord3index1 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 1:6) %>% 
  as.data.frame()

ord3index2 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 7:12) %>% 
  as.data.frame()

ord3index3 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 13:18) %>% 
  as.data.frame()
```

```{r}
brayByPond <- c()
for (i in 1:length(ord3index1$V1)){
  brayByPond <- rbind(brayByPond, 
                       c("T106", brayCurtisInv[ord3index1$V1[i], ord3index1$V2[i]]))
}


for (i in 1:length(ord3index2$V1)){
  brayByPond <- rbind(brayByPond, 
                       c("T108", brayCurtisInv[ord3index2$V1[i], ord3index2$V2[i]]))
}


for (i in 1:length(ord3index3$V1)){
  brayByPond <- rbind(brayByPond, 
                       c("T109", brayCurtisInv[ord3index3$V1[i], ord3index3$V2[i]]))
}

brayByPond <- as.data.frame(brayByPond)
colnames(brayByPond) <- c("pond", "bcSim")
brayByPond$bcSim <- as.numeric(brayByPond$bcSim)
```

```{r}
brayByPond %>% 
  ggplot(aes(x = pond, y = bcSim, group = pond)) +
  geom_boxplot() + 
  labs(x = "Pond", y = "Similarity (Bray-Curtis)", title = "Bray-Curtis Similarity versus Pond")
```


# For real


## Water Chemistry

```{r}
tempModel <- lmer(Temperature ~ Depth*Num + (1|Pond), data = verticalSubData, REML = FALSE)
summary(tempModel)
```


```{r}
doModel <- lmer(DO ~ Depth*Num + (1|Pond), data = verticalSubData, REML = FALSE)
summary(doModel)
```


```{r}
condModel <- lmer(Conductivity ~ Depth*Num + (1|Pond), data = verticalSubData, REML = FALSE)
summary(condModel)
```


```{r}
turbModel <- lmer(Turbidity ~ Depth*Num + (1|Pond), data = verticalSubData, REML = FALSE)
summary(turbModel)

#rand.slope.turb.model<-lmer(Turbidity~Depth*Num +(1+Depth|Pond), data = verticalSubData) #abject Failure
```

```{r}
temp <- verticalSubData %>% 
  group_by(Pond, Depth) %>%
  filter(Depth %in% c(50, 100, 150)) %>% 
  select(Temperature, DO, Conductivity, Turbidity) %>% 
  summarize(meanTemp = mean(Temperature), meanDO = mean(DO), meanCond = mean(Conductivity), 
            meanTurb = mean(Turbidity))

temp1 <- rbind(temp, data.frame(Pond = "T109", Depth = 150, 
                    meanTemp = NA,
                    meanDO = NA, 
                    meanCond = NA, 
                    meanTurb = NA))

temp2 <- rbind(temp, data.frame(Pond = "T109", Depth = 150, 
                    meanTemp = NA,
                    meanDO = NA, 
                    meanCond = NA, 
                    meanTurb = NA))

sites <- c(rep("A", 9), rep("B", 9))

dfMeanWater <- cbind(rbind(temp1, temp2), sites = sites)
dfMeanWater
```



## Diversity Measures

```{r}
numericalInvData <- invertData[,-c(1,2)]
SR <- specnumber(numericalInvData)
```

```{r Sim}
simp.inv <- diversity(numericalInvData, index = "invsimpson")
simp.even <- simp.inv / SR
simp.even
```

```{r Shn}
shan <- diversity(numericalInvData, index = "shannon")
exp.shan <- exp(shan)
shan.even <- shan / (log(SR))
dfDivFull <- cbind(dfMeanWater, data.frame(SR, simp.inv, simp.even, shan, shan.even, exp.shan))
dfDiv <- dfDivFull[-c(9, 18),]

dfDivFull
```


```{r}
dfDivFull %>% 
  group_by(Pond) %>% 
  summarize(meanSR = mean(SR),
           meanSimp = mean(simp.inv),
           meanSimpEv = mean(simp.even),
           meanShan = mean(shan),
           meanShanEv = mean(shan.even),
           meanShanEx = mean(exp.shan))

dfDivFull %>% 
  group_by(Depth) %>% 
  summarize(meanSR = mean(SR),
           meanSimp = mean(simp.inv),
           meanSimpEv = mean(simp.even),
           meanShan = mean(shan),
           meanShanEv = mean(shan.even),
           meanShanEx = mean(exp.shan))

mean(dfDivFull$SR)
mean(dfDivFull$simp.inv)
mean(dfDivFull$simp.even)
mean(dfDivFull$shan)
mean(dfDivFull$shan.even)
mean(dfDivFull$exp.shan)
```



```{r Rare}
rarecurve(numericalInvData)
community.N <- rowSums(numericalInvData)
smallest.N <- min(community.N)
lines(x = c(smallest.N, smallest.N), y = c(0, max(SR)))
```




```{r SR}
srModel <- lmer(SR ~ Depth + (1|Pond) + (sites|Pond), 
                data = dfDiv,
                REML = FALSE)
summary(srModel)


srModel <- lm(SR ~ Depth*Pond, 
                data = dfDiv)
summary(srModel)
Anova(srModel, type = "III")
```

```{r}
dfDivFull %>% 
  ggplot(aes(x = as.character(Depth), y = SR, colour = Pond, shape = factor(sites))) +
  geom_jitter()  +
  scale_x_discrete(limits = c("50", "100", "150")) + 
  labs(title = "Species Richness versus Depth", x = "Depth (cm)", y = "Species Richness",
       colour = "Pond ID", shape = "Site")
ggsave("Plots/sr.png")


dfDivFull %>% 
  ggplot(aes(x = as.character(Depth), y = exp.shan, colour = Pond, shape = factor(sites))) +
  geom_jitter()  +
  scale_x_discrete(limits = c("50", "100", "150")) + 
  labs(title = "Exponential Shannon Index versus Depth", x = "Depth (cm)", y = "Exponential Shannon Index",
       colour = "Pond ID", shape = "Site")
ggsave("Plots/shan.png")


dfDivFull %>% 
  ggplot(aes(x = as.character(Depth), y = shan.even, colour = Pond, shape = factor(sites))) +
  geom_jitter()  +
  scale_x_discrete(limits = c("50", "100", "150")) + 
  labs(title = "Shannon Evenness versus Depth", x = "Depth (cm)", y = "Shannon Evenness",
       colour = "Pond ID", shape = "Site")
ggsave("Plots/shanev.png")
```


```{r Simp}
simpModel <- lmer(simp.inv ~ Depth + (1|Pond) + (sites|Pond), 
                data = dfDiv,
                REML = FALSE)
summary(simpModel)
```


```{r SimpEv}
simpEvModel <- lmer(simp.even ~ Depth + (1|Pond) + (sites|Pond), 
                data = dfDiv,
                REML = FALSE)
summary(simpEvModel)
```


```{r Shan}
shanModel <- lmer(exp.shan ~ Depth + (1|Pond) + (sites|Pond), 
                data = dfDiv,
                REML = FALSE)
summary(shanModel)

shanModel.new <- lm(exp.shan ~ Pond, 
                data = dfDiv)
summary(shanModel.new)
Anova(shanModel.new, type = "III")
```


```{r ShanEv}
shanEvModel <- lmer(shan.even ~ Depth + (1|Pond) + (sites|Pond), 
                data = dfDiv,
                REML = FALSE)
summary(shanEvModel)


shanEvModel.new <- lm(shan.even ~ Depth*Pond, 
                data = dfDiv)
summary(shanEvModel.new)
Anova(shanEvModel.new, type = "III")
```

```{r}

shanEvModel.new <- lm(shan.even ~ Depth*Pond, 
                data = dfDiv)
summary(shanEvModel.new)
Anova(shanEvModel.new, type = "III")
```


## Similarity 

```{r}
brayCurtisInv <- 1 - vegdist(numericalInvData, method = "bray", diag = FALSE) %>% 
  as.matrix()

bray.matrix<-vegdist(numericalInvData, method = "bray")
#sorr

#### Making into relative abundances 
Manipulating.wide<-numericalInvData
Manipulating.wide$Identifier<-rownames(numericalInvData)
#Calculating sample size
sample.size.vec<-as.vector(rowSums(numericalInvData))
```
#Model predicting total abujndance
```{r}
dfDivFull$Sample.Size<-c(202, 144, 57, 179, 110, 66, 627, 132, 191, 331, 134, 138, 80, 78, 16, 281, 205, 180)
## 
s.s.lm<-lm(Sample.Size~Depth,data=dfDivFull)
summary(s.s.lm)
Anova(s.s.lm, type = "III")

dfDivFull %>% 
  ggplot(aes(x = as.character(Depth), y = Sample.Size, colour = Pond, shape = factor(sites))) +
  geom_jitter()  +
  scale_x_discrete(limits = c("50", "100", "150")) + 
  labs(title = "Sample Size versus Depth", x = "Depth (cm)", y = "Sample Size",
       colour = "Pond ID", shape = "Site")
ggsave("Plots/ss.png")
```


### Bray Curtis By Depth

```{r}
mod3index1 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(1, 18, 3)) %>% 
  as.data.frame()

mod3index2 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(2, 18, 3)) %>% 
  as.data.frame()

mod3index3 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(3, 18, 3)) %>% 
  as.data.frame()
```


```{r}
brayByDepth <- c()
for (i in 1:length(mod3index1$V1)){
  brayByDepth <- rbind(brayByDepth, 
                       c(50, brayCurtisInv[mod3index1$V1[i], mod3index1$V2[i]]))
}


temp2 <- 0 
for (i in 1:length(mod3index2$V1)){
  brayByDepth <- rbind(brayByDepth, 
                       c(100, brayCurtisInv[mod3index2$V1[i], mod3index2$V2[i]]))
}


temp3 <- 0 
for (i in 1:length(mod3index3$V1)){
  brayByDepth <- rbind(brayByDepth, 
                       c(150, brayCurtisInv[mod3index3$V1[i], mod3index3$V2[i]]))
}

brayByDepth <- as.data.frame(brayByDepth)
colnames(brayByDepth) <- c("depth", "bcSim")
```


```{r}
brayByDepth %>% 
  ggplot(aes(x = depth, y = bcSim, group = depth)) +
  geom_boxplot() + 
  labs(x = "Depth (cm)", y = "Similarity (Bray-Curtis)", title = "Bray-Curtis Similarity versus Depth")
ggsave("Plots/bcD.png")
```


### Bray-Curtis By Pond


```{r}
ord3index1 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 1:6) %>% 
  as.data.frame()

ord3index2 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 7:12) %>% 
  as.data.frame()

ord3index3 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 13:18) %>% 
  as.data.frame()
```

```{r}
brayByPond <- c()
for (i in 1:length(ord3index1$V1)){
  brayByPond <- rbind(brayByPond, 
                       c("T106", brayCurtisInv[ord3index1$V1[i], ord3index1$V2[i]]))
}


for (i in 1:length(ord3index2$V1)){
  brayByPond <- rbind(brayByPond, 
                       c("T108", brayCurtisInv[ord3index2$V1[i], ord3index2$V2[i]]))
}


for (i in 1:length(ord3index3$V1)){
  brayByPond <- rbind(brayByPond, 
                       c("T109", brayCurtisInv[ord3index3$V1[i], ord3index3$V2[i]]))
}

brayByPond <- as.data.frame(brayByPond)
colnames(brayByPond) <- c("pond", "bcSim")
brayByPond$bcSim <- as.numeric(brayByPond$bcSim)
```

```{r}
brayByPond %>% 
  ggplot(aes(x = pond, y = bcSim, group = pond)) +
  geom_boxplot() + 
  labs(x = "Pond", y = "Similarity (Bray-Curtis)", title = "Bray-Curtis Similarity versus Pond")
ggsave("Plots/bcP.png")
```


```{r}
paInvData <- numericalInvData
paInvData[paInvData > 0] <- 1

sorensonInv <- 1 - vegdist(numericalInvData, method = "jaccard", diag = FALSE) %>% 
  as.matrix()
```


### Jaccard By Depth

```{r}
mod3index1 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(1, 18, 3)) %>% 
  as.data.frame()

mod3index2 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(2, 18, 3)) %>% 
  as.data.frame()

mod3index3 <- combinations(n = 6, r = 2, repeats.allowed = F, v = seq(3, 18, 3)) %>% 
  as.data.frame()
```


```{r}
sorenByDepth <- c()
for (i in 1:length(mod3index1$V1)){
  sorenByDepth <- rbind(sorenByDepth, 
                       c(50, sorensonInv[mod3index1$V1[i], mod3index1$V2[i]]))
}


temp2 <- 0 
for (i in 1:length(mod3index2$V1)){
  sorenByDepth <- rbind(sorenByDepth, 
                       c(100, sorensonInv[mod3index2$V1[i], mod3index2$V2[i]]))
}


temp3 <- 0 
for (i in 1:length(mod3index3$V1)){
  sorenByDepth <- rbind(sorenByDepth, 
                       c(150, sorensonInv[mod3index3$V1[i], mod3index3$V2[i]]))
}

sorenByDepth <- as.data.frame(sorenByDepth)
colnames(sorenByDepth) <- c("depth", "sorSim")
```


```{r}
sorenByDepth %>% 
  ggplot(aes(x = depth, y = sorSim, group = depth)) +
  geom_boxplot() + 
  labs(x = "Depth (cm)", y = "Similarity (Jaccard)", title = "Jaccard Similarity versus Depth")
ggsave("Plots/jacD.png")
```


### Sorenson By Pond


```{r}
ord3index1 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 1:6) %>% 
  as.data.frame()

ord3index2 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 7:12) %>% 
  as.data.frame()

ord3index3 <- combinations(n = 6, r = 2, repeats.allowed = F, v = 13:18) %>% 
  as.data.frame()
```

```{r}
sorenByPond <- c()
for (i in 1:length(ord3index1$V1)){
  sorenByPond <- rbind(sorenByPond, 
                       c("T106", sorensonInv[ord3index1$V1[i], ord3index1$V2[i]]))
}


for (i in 1:length(ord3index2$V1)){
  sorenByPond <- rbind(sorenByPond, 
                       c("T108", sorensonInv[ord3index2$V1[i], ord3index2$V2[i]]))
}


for (i in 1:length(ord3index3$V1)){
  sorenByPond <- rbind(sorenByPond, 
                       c("T109", sorensonInv[ord3index3$V1[i], ord3index3$V2[i]]))
}

sorenByPond <- as.data.frame(sorenByPond)
colnames(sorenByPond) <- c("pond", "sorSim")
sorenByPond$sorSim <- as.numeric(sorenByPond$sorSim)
```

```{r}
sorenByPond %>% 
  ggplot(aes(x = pond, y = sorSim, group = pond)) +
  geom_boxplot() + 
  labs(x = "Pond", y = "Similarity (Jaccard)", title = "Jaccard Similarity versus Pond")
ggsave("Plots/jacP.png")
```

```{r}
brayCurtisDis <- vegdist(numericalInvData[-c(9,18),], method = "bray")

adonis2(brayCurtisDis ~ Pond + Depth + meanTemp + meanDO + meanCond + meanTurb,
        data = dfDiv)

bc.pcoa<-cmdscale(brayCurtisDis,eig = T)
ordiplot(bc.pcoa)
#how well do these axes fit
bc.pcoa$eig/sum(bc.pcoa$eig)
```






```{r}
Long.data<-invertData %>% 
  filter(pondID != "T109A3" & pondID != "T109B3") %>% 
  pivot_longer(!c(pondID, depth), values_to = "total", names_to = "family")

#getting sample sizes
Long.data<-Long.data %>% group_by(pondID,depth) %>% mutate(Sample_Size=sum(total)) %>% ungroup()
Long.data<-Long.data %>% group_by(pondID,depth) %>% mutate(Relative.Abundance=total/Sample_Size) %>% ungroup()

#Flip it back into a wide format using relative abundances as the values 

wide.relabund <- pivot_wider(Long.data,id_cols = c(pondID, depth),  names_from = "family", values_from = Relative.Abundance)

#remove the depth column 
wide.relabund<-wide.relabund %>% dplyr::select(-depth)
wide.relabund<-column_to_rownames(wide.relabund,var="pondID")

#### Getting distance matrixes from relative abundances 
Rel.BC<-vegdist(wide.relabund,method="bray")
Rel.Sorr<-vegdist(wide.relabund,method="jaccard")

```


#### Visualizing similarity among samples ####

```{r}
bc.pcoa<-cmdscale(Rel.BC,eig = T)
ordiplot(bc.pcoa)
#how well do these axes fit
bc.pcoa$eig/sum(bc.pcoa$eig) #78% of variation explained in axis 1, 11% in axis 2
# They fit really well 


####### 
Jacc.pcoa<-cmdscale(Rel.Sorr,eig = T)
ordiplot(Jacc.pcoa)
Jacc.pcoa$eig/sum(Jacc.pcoa$eig) #Similarly good 65% in first and 16% in second 80% in first 2 axes 



#####
#Let's extract the axes to plot 
BC.PCOA.axis1<-bc.pcoa$points[1:16]
BC.PCOA.axis2<-bc.pcoa$points[17:32]

#Jaccard
JC.PCOA.axis1<-Jacc.pcoa$points[1:16]
JC.PCOA.axis2<-Jacc.pcoa$points[17:32]

####
#Gluing these back onto a dataframe with identifying information 


POND <- c("T106A1", "T106A2", "T106A3",
          "T108A1", "T108A2", "T108A3",
          "T109A1", "T109A2",
          "T106B1", "T106B2", "T106B3",
          "T108B1", "T108B2", "T108B3",
          "T109B1", "T109B2")

dfDiv <- cbind(dfDiv, pondID = POND)
## Gluing 
dfDiv$BC.PCOA.axis1<-BC.PCOA.axis1
dfDiv$BC.PCOA.axis2<-BC.PCOA.axis2
dfDiv$JC.PCOA.axis1<-JC.PCOA.axis1
dfDiv$JC.PCOA.axis2<-JC.PCOA.axis2

# PLotting pretty

#Do we see greater clustering by pond or by depth? 
dfDiv %>% 
  ggplot(aes(x=BC.PCOA.axis1,BC.PCOA.axis2,
             colour=as.character(Depth),shape=Pond, group =Depth)) +
  geom_point() +
  stat_ellipse() + 
  labs(title = "PERMANOVA (Bray Curtis, Depth Cluster)", x = "Bray Curtis Axis 1", y = "Bray Curtis Axis 2",
       shape = "Pond ID", colour = "Depth (cm)")
ggsave("Plots/clusterBCd.png")


dfDiv %>% ggplot(aes(x=BC.PCOA.axis1,BC.PCOA.axis2,color=as.character(Pond),
                     shape=as.character(Depth), group =Pond)) + 
  geom_point() + 
  stat_ellipse() + 
  labs(title = "PERMANOVA (Bray Curtis, Pond Cluster)", x = "Bray Curtis Axis 1", y = "Bray Curtis Axis 2",
       shape = "Depth (cm)", colour = "Pond ID")
ggsave("Plots/clusterBCp.png")


##### Jaccard
dfDiv %>% ggplot(aes(x=JC.PCOA.axis1,JC.PCOA.axis2,color=as.character(Depth),
                     shape=Pond, group =Depth)) + 
  geom_point() +
  stat_ellipse() + 
  labs(title = "PERMANOVA (Jaccard, Depth Cluster)", x = "Jaccard Axis 1", y = "Jaccard Axis 2",
       shape = "Pond ID", colour = "Depth (cm)")
ggsave("Plots/clusterJCd.png")


dfDiv %>% ggplot(aes(x=JC.PCOA.axis1,JC.PCOA.axis2,color=as.character(Pond),
                     shape=as.character(Depth), group =Pond)) +
  geom_point() +
  stat_ellipse() + 
  labs(title = "PERMANOVA (Jaccard, Pond Cluster)", x = "Jaccard Axis 1", y = "Jaccard Curtis Axis 2",
       shape = "Depth (cm)", colour = "Pond ID")
ggsave("Plots/clusterJCp.png")

#Visually it looks like deeper are a subset of shallower. Can we see if any of our chemistry or depth parameters are predicting this? 
```

### What's driving patterns?

```{r}
### First perform model selection with distance-based RDA (db-RDA) then permanova() of the significant factors 

#We need to scale our variables to common units 
dfDiv<-dfDiv %>% ungroup()
Things.to.scale<-dfDiv %>% dplyr::select(pondID,meanTemp,meanDO,meanCond,meanTurb)
scaler.int<-scale(Things.to.scale[,2:5])
scaler.int.df<-as.data.frame(scaler.int)
# Extracting as vectors and pasting back on 
dfDiv$Z.Temp<-as.vector(scaler.int.df$meanTemp)
dfDiv$Z.DO<-as.vector(scaler.int.df$meanDO)
dfDiv$Z.Cond<-as.vector(scaler.int.df$meanCond)
dfDiv$Z.Turb<-as.vector(scaler.int.df$meanTurb)

#### making a categorical depth
dfDiv$Cat.depth<-as.character(dfDiv$Depth)
#
BC.db_RDA<-capscale(Rel.BC~Depth+Z.Temp+Z.DO+Z.Cond+Z.Turb,data=dfDiv)
ordistep(BC.db_RDA) #Just DO seems important

cor(dfDiv$Depth,dfDiv$Z.DO) #Nearly perfectly correlated so hard to say whether driven by DO or some other ocrrelate with depth 

#### JAcc

JC.db_RDA<-capscale(Rel.Sorr~Depth+Z.Temp+Z.DO+Z.Cond+Z.Turb,data=dfDiv)
ordistep(JC.db_RDA) #Just DO 

#### Anything different with a permanova?

perm.1<-adonis2(Rel.BC~Depth+Z.Temp+Z.DO+Z.Cond+Z.Turb,data=dfDiv,permutations=9999)

perm.2<-adonis2(Rel.Sorr~Depth+Z.Temp+Z.DO+Z.Cond+Z.Turb,data=dfDiv,permutations=99999)
#Nothing; marginally non-significant effect of DO

### Early results at least suggest that at the community level, only Dissolved oxygen predicts differences in biological communities. and dissolved oxygen changes with depth (Decreases)



##### DO we have sig effect of temperature?

```



### Univariate Modelling ## 

General approach of not using a random effects model since sample size is low and then using the step function 
```{r}
Saturated.Shann.div<-lm(shan~Depth + Z.Temp + Z.DO + Z.Cond + Z.Turb+Pond, data = dfDiv)
step(Saturated.Shann.div)
summary(Saturated.Shann.div)
Anova(Saturated.Shann.div,type="III")

#Selected chemistry 
selected.shannon.div.lm<-lm(shan~Depth + Z.Temp + Z.DO + Z.Cond +Pond, data = dfDiv)
summary(selected.shannon.div.lm)
Anova(selected.shannon.div.lm, type = "III")
# None of our chemistry variables significantly predict shannon diversity
```
```{r}
Saturated.SR.div<-lm(SR~Depth + Z.Temp + Z.DO + Z.Cond + Z.Turb+Pond, data = dfDiv)
step(Saturated.SR.div)

selected.SR.div<-lm(SR~Depth + Z.Temp + Z.DO + Z.Cond+Pond, data = dfDiv)
summary(selected.SR.div)
Anova(selected.SR.div,type="III")

#Selected chemistry 
selected.shannon.div.lm<-lm(shan~Z.DO ,data=dfDiv)
summary(selected.shannon.div.lm)
anova(selected.shannon.div.lm)
# None of our chemistry variables significantly predict shannon diversity


#Does model output change iwth depfth as numeric? 
aturated.SR.div<-lm(SR~Cat.depth + Z.Temp + Z.DO + Z.Cond + Z.Turb+Pond, data = dfDiv)
step(aturated.SR.div)
selected.SR.div.lm<-lm(SR ~ Cat.depth + Z.Cond + Pond, data = dfDiv)
summary(selected.SR.div.lm)
Anova(selected.SR.div.lm,type="III")


#shan with cat depth?? 

aturated.shan.div<-lm(shan~Cat.depth + Z.Temp + Z.DO + Z.Cond + Z.Turb+Pond, data = dfDiv)
step(aturated.shan.div)
selected.shan.div.lm<-lm(shan ~ Cat.depth + Z.Cond + Pond, data = dfDiv)
summary(selected.shan.div.lm)
Anova(selected.shan.div.lm,type="III")

aturated.shan.div<-lm(shan~Z.Temp*Cat.depth + Z.DO*Cat.depth + Z.Cond*Cat.depth + Z.Turb, data = dfDiv)
step(aturated.shan.div)
selected.shan.div.lm<-lm(shan ~ Z.Temp + Cat.depth + Z.DO + Z.Cond + Z.Temp:Cat.depth, data = dfDiv)
summary(selected.shan.div.lm)
Anova(selected.shan.div.lm,type="III")
```

```{r}
selected.shan.div.lm<-lm(shan ~ Z.Cond*Depth, data = dfDiv)
summary(selected.shan.div.lm)
Anova(selected.shan.div.lm,type="III")

selected.shan.div.lm<-lm(SR ~ Z.Cond*Depth, data = dfDiv)
summary(selected.shan.div.lm)
Anova(selected.shan.div.lm,type="III")
```


```{r}
dfDiv %>% 
  ggplot(aes(x = Z.Cond, y = shan, colour = as.character(Depth))) +
  geom_jitter()  +
  labs(title = "Shannon Index versus Conductivity", x = "Conductivity (Standardized)", y = "Shannon Index",
       colour = "Depth (cm)")
ggsave("Plots/condShan.png")

dfDiv %>% 
  ggplot(aes(x = Z.Cond, y = SR, colour = as.character(Depth))) +
  geom_jitter()  +
  labs(title = "Species Richness versus Conductivity", x = "Conductivity (Standardized)", y = "Species Richness",
       colour = "Depth (cm)")
ggsave("Plots/condSR.png")
```



```{r}
cor(dfDiv[,19:22])
cor(dfDiv$meanTemp, dfDiv$meanDO, )

### how does chem cluster with pca? 
Stanrdard.chem.pca<-prcomp(dfDiv[,19:22])
biplot(Stanrdard.chem.pca)
#Loadings
Stanrdard.chem.pca$rotation
```


